FlowFPX: Nimble Tools for Debugging Floating-Point Exceptions

### 2023-08-04 paper.tex
Cleanup bibliography
Rough draft of section 3 complete!

%% --- autogenerated TOC ---
% juliacon
% document
%   Introduction
%   Floating Point Exception Primer
%     Where do \NaN{
%       TODO good quote, try to work this in
%     Infinity
%     To Kill a Floating Point
%   \FlowFPX{
%     \FT{
%       Tracking Excptional Values
%       Fuzzing
%     Visualization
%     \GPUFPX{
%   Case Studies
%     Surprise \NaN{
%       CSTGs help us get a bigger picture
%     \NaN{
%     Finch
%     A Bug-Hunting Safari
%     \GPUFPX{
%   Related Work
%   Discussion
%     Tracking more than \NaN{
%     Enhanced Fuzzing
%     Static FPX, Gradual Types
%   Acknowledgments
% document
%% --- end autogenerated TOC ---

@@

Coalesced Stack Trace Graphs, or CSTGs\cite{humphreySystematicDebuggingMethods2014} provide a way to visualize large amounts of stack traces in a compact form.
\FT{} produces

@@

FILL basic idea, loose coupling, typical overhead,

@@

\subsection{Surprise \NaN{}s: ShallowWaters}

@@

\begin{verbatim}
…
-                 FT/TrackedFloat.jl:106
momentum_u!       ShallowWaters/rhs.jl:246
rhs_nonlinear!    ShallowWaters/rhs.jl:50
rhs!              ShallowWaters/rhs.jl:14 [inlined]
time_integration  ShallowWaters/time_integration.jl:77
run_model         ShallowWaters/run_model.jl:37
…
\end{verbatim}

@@

\subsection{\GPUFPX{}: Oceananigans}

@@

%% Demmel, somewhere!
%% https://arxiv.org/pdf/2207.09281.pdf

@@

FPChecker, " considers CPU OpenMP and MPI codes, again using LLVM
instrumentation"~\cite{ltlg-iiswc-2022}

@@

CSTG~\cite{hmbdrg-cse-2014}.
What else???

@@

In addition to the cost incurred by intercepting \fp{} operations, gathering stack traces is expensive---we measured a significant slowdown on the order of 10x relative to a non-logging setup of \FT{}.
We recommend that users of \FT{} make use of the \texttt{maxLogs} and \texttt{exclude_stacktrace} configuration options to limit the number and kind of logs gathered, and thereby reduce the number of calls to \texttt{stacktrace()}.

>>>>

%% --- autogenerated TOC ---
% juliacon
% document
%   Introduction
%   Floating Point Exception Primer
%     Where do \NaN
%       TODO good quote, try to work this in
%     Infinity
%     To Kill a Floating Point
%   \FlowFPX
%     \FT
%       Tracking Excptional Values
%       Fuzzing
%     Visualization
%     \GPUFPX
%   Case Studies
%     Surprise \NaN
%       CSTGs help us get a bigger picture
%     \NaN
%     Finch
%     A Bug-Hunting Safari
%     \GPUFPX
%   Related Work
%   Discussion
%     Tracking more than \NaN
%     Enhanced Fuzzing
%     Static FPX, Gradual Types
%   Acknowledgments
% document
%% --- end autogenerated TOC ---

@@

\FT{} can produce copious amounts of log files which can be challenging to sift through manually.
Coalesced Stack Trace Graphs, or CSTGs\cite{humphreySystematicDebuggingMethods2014} provide a way to visualize large amounts of stack traces in a compact form.
CSTG is a stand-alone tool and was not designed with \FT{} specifically in mind.
However, it pairs well with \FT{} and makes coming through the log files easier.

CSTG takes a set of stack traces and makes a graph: for each stack trace, it creates a node for every frame, and then creates or increments an edge between two adjacent frames, as seen in \cref{fig:cstg_demo}.

\begin{figure}
  \centering
  \includegraphics[width=3in]{./fig/cstg_static_diagram.png}
  \caption{CSTG makes a graph by connecting adjacent stack frames}
  \label{fig:cstg_demo}
\end{figure}

CSTG's output looks like the output seen in \cref{fig:sw_nan_cstg}, which was generated by running \FT{} on a library called ShallowWaters.
We will go into more detail about ShallowWaters in \cref{s:sw}.
This graph has been edited for clarity and to fit into the size constraints of this paper.

These graphs are meant to be read bottom-up: the bottom nodes are the roots where exceptional values manifested.
Higher nodes in the graph show function calls progressively closer to the top-level call that spawned them.
Thick lines and high numbers indicate flows with the highest volume leading to errors---though thin areas are still worth investigating.
CSTG provides a map to aid in contextualizing a stack trace in a log file.

\subsubsection{Graph diffing}

CSTG can diff two graphs to show how the flows changed.
For example, we took the \FT{} logs from running ShallowWaters\cite{klowerNumberFormatsError2020} \footnote{\Cref{s:sw}}, and split the logs at the 10\% mark.
We used CSTG to diff the flow graphs between the first 10\% of the run of the program and the remaining 90\%.
The diff can be seen in \cref{fig:cstg_diff_demo}.

\begin{figure}
  \centering
  \includegraphics[width=3in]{./fig/cstg_diff_pretty.png}
  \caption{Diff between the early and latter stages of ShallowWaters}
  \label{fig:cstg_diff_demo}
\end{figure}

The green lines indicate flows that were \emph{new} in the latter 90\% of the program's run.
The red lines indicate flows that were \emph{only in} the first 10\% of the program's run.
What this means depends on the particular program; a domain expert can use these clues to e.g. find where instability started in the first part of the program, and where that instability got reinforced later on.

@@

% Acknowledge that apps get offloaded to GPU; FT doesn't do that, but there's this companion tool…
% You run it on a kernel; catches things not caught by GPUs; see Oceananigans as an example below
% TODO: add note in future work/discussion about this as well; other languages too—not just GPU
% FloatTracker could supply an API that wraps calls to a C library—lib is still a black box, but FT gives you a big picture

Many programs offload work onto GPUs, which are no less susceptible to \fp{} exceptions than CPUs.
In some regards they are \emph{more} susceptible because GPUs do not have the same exception-handling mechanisms that CPUs do.
Thus, tools to catch \fp{} exceptions on the GPU are essential.
\FT{} is constrained to CPU workloads; however, there is a companion tool called \GPUFPX{}\cite{llsflg-hpdc-2023} that can help diagnose \fp{} exceptions on the GPU.
\GPUFPX{} can detect floating-point exceptions on the GPU that GPU code cannot do on its own.
\GPUFPX{} reports these exceptions back to the CPU for further analysis.

\FT{} and \GPUFPX{} can cover the entire computational stack, as seen in \cref{fig:ft_and_gpu_fpx}.
Work remains to tighten the integration between \FT{} and \GPUFPX{} to reduce manual work to employ both tools for heterogeneous computation.

\begin{figure}
  \centering
  \includegraphics[width=3in]{./fig/float_tracker_context_boxes.png}
  \caption{FloatTracker and GPU-FPX compliment each other}
  \label{fig:ft_and_gpu_fpx}
\end{figure}

See \cref{s:ocean} for an example of \GPUFPX{} in action, and \ref{f:gpufpx} for an example of its output.

@@

\subsection{Surprise \NaN{}s: ShallowWaters}
\label{s:sw}

@@

\begin{verbatim}
-                 FT/TrackedFloat.jl:106
momentum_u!       ShallowWaters/rhs.jl:246
rhs_nonlinear!    ShallowWaters/rhs.jl:50
rhs!              ShallowWaters/rhs.jl:14 [inlined]
time_integration  ShallowWaters/time_integration.jl:77
run_model         ShallowWaters/run_model.jl:37
top-level
\end{verbatim}

@@

\subsection{\GPUFPX{}: Oceananigans}
\label{s:ocean}

@@

Demmel, somewhere!
%% https://arxiv.org/pdf/2207.09281.pdf

@@

FPChecker, " considers CPU OpenMP and MPI codes, again using LLVM
instrumentation"~\cite{ltlg-iiswc-2022}
% FPChecker has a nice table of tools; some already mentioned here

@@

CSTG~\cite{humphreySystematicDebuggingMethods2014}.
What else???

@@

In addition to the cost incurred by intercepting \fp{} operations, gathering stack traces is expensive---we measured a significant slowdown on the order of 10x relative to a non-logging setup of \FT{}.
We recommend that users of \FT{} make use of the \texttt{maxLogs} and \texttt{exclude\_stacktrace} configuration options to limit the number and kind of logs gathered, and thereby reduce the number of calls to \texttt{stacktrace()}.
